---
title: "Iris"
author: "Bill"
date: "December 17, 2018"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(mlbench)
library(e1071)
library(corrplot)
library(rcompanion)
library(klaR)
library(MASS)
library(stats)
```

## Data Introduction
```{r echo=FALSE, comment=NA}
# load and review the data basics
```
### Snapshot of Header Data
```{r echo=FALSE, comment=NA}
data("iris")
dataset <- iris
head(dataset)
```
### Data Observations, Number of Variables, Classifications
```{r echo=FALSE, comment=NA}
dim(dataset)
sapply(dataset,class)
levels(dataset$Species)
```
### Standard Deviation for Continuous Variables
```{r echo=FALSE, comment=NA}
sapply(dataset[,1:4],sd)
```
### Basic Statistics
```{r echo=FALSE, comment=NA}
summary(dataset)
```
### Skewness and Correlations
```{r echo=FALSE, comment=NA}
skew <- apply(dataset[,1:4],2,skewness)
print(skew)
correlations <- cor(dataset[,1:4])
print(correlations)
corrplot(correlations, method = "circle")
```

### Basic Raw Data Graphical Analysis
```{r echo=FALSE, comment=NA}
# create histograms for each attribute
par(mfrow = c(1,4))
for (i in 1:4) {
  hist(dataset[,i], main = names(dataset)[i])
}
# create a layout of simpler density plots by attribute
par(mfrow = c(1,4))
for (i in 1:4) {
  plot(density(dataset[,i]), main = names(dataset)[i])
}
# density plots for each attribute by class value
x <- dataset[,1:4]
y <- dataset[,5]
scales <- list(x = list(relation = "free"), y = list(relation = "free"))
featurePlot(x = x, y = y, plot = "density", scales = scales)
# Create separate boxplots for each attribute
par(mfrow = c(1,4))
for (i in 1:4) {
  boxplot(dataset[,i], main = names(dataset)[i])
}
# box and whisker plots for each attribute by class value
x <- dataset[,1:4]
y <- dataset[,5]
featurePlot(x = x, y = y, plot = "box")
# pairwise plot
pairs(dataset)
# pairwise scatter plots colored by class
pairs(Species~., data = dataset, col = dataset$Species)
# scatter plot matrix
featurePlot(x = x, y = y, plot = "ellipse")
```

## Prediction
```{r echo=FALSE, comment=NA}
# create a validation dataset
validationIndex <- createDataPartition(dataset$Species, p = 0.80, list = FALSE)
validation <- dataset[-validationIndex,]  # select 20% of the data for validation
dataset <- dataset[validationIndex,]  # select 80% of the data for training
# generate a test harness running algorithms using 10-fold cross validation
trainControl <- trainControl(method = "cv", number = 10)
metric <- "Accuracy"  # ratio of number of correctly predicted instances divided by the total instances
# 5 candidate models
#   Linear Discriminant Analysis (LDA)
#   Classification and Regression Trees (CART)
#   k-Nearest Neighbors (KNN)
#   Support Vector Machines (SVM) with a radial kernal
#   Random Forest (RF)
```

### Models to Try:
#### Linear Discriminant Analysis (LDA)
#### Classification and Regression Trees (CART)
#### k-Nearest Neighbors (KNN)
#### Support Vector Machines (SVM) with a radial kernal
#### Random Forest (RF)

```{r echo=FALSE, comment=NA}
set.seed(7)
fit.lda <- train(Species~., data = dataset, method = "lda", metric = metric, trControl = trainControl)
set.seed(7)
fit.cart <- train(Species~., data = dataset, method = "rpart", metric = metric, trControl = trainControl)
set.seed(7)
fit.knn <- train(Species~., data = dataset, method = "knn", metric = metric, trControl = trainControl)
set.seed(7)
fit.svm <- train(Species~., data = dataset, method = "svmRadial", metric = metric, trControl = trainControl)
set.seed(7)
fit.rf <- train(Species~., data = dataset, method = "rf", metric = metric, trControl = trainControl)
results <- resamples(list(lda = fit.lda, cart = fit.cart, knn = fit.knn, svm = fit.svm, rf = fit.rf))
summary(results)
dotplot(results)
print(fit.lda)
# estimate skill of LDA on the validation dataset
predictionsLDA <- predict(fit.lda, validation)
confusionMatrix(predictionsLDA, validation$Species)
```